---
title: "Support for dplyr"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Support for dplyr}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup, message=FALSE, warning=FALSE}
set.seed(123)
library(strapgod)
library(dplyr)
```

# Introduction

As much as possible, strapgod attempts to let you use any dplyr function that you want on the `resampled_df` object that is returned by `bootstrapify()` and `samplify()`. Some functions have specialized behavior, like `summarise()`, while most others just call `collect()` to materialize the bootstrap rows before passing on to the underlying dplyr function.

What follows is a list of the dplyr functions that have "special" properties when used on a `resampled_df`.

# collect()

The most important dplyr function for strapgod is `collect()`. Generally, this has been used to force a computation from a data base query and return the results as a tibble, and it has a similar context here. `collect()` forces the materialization of the virtual groups, and returns the full grouped tibble back to you.

```{r}
x <- bootstrapify(iris, 10)

# Not materialized
x

# Materialized
collect(x)
```

When calling `collect()` directly, there are two arguments available to extract extra information about the bootstraps. 

`id` adds a sequence of integers from `1:n` for each bootstrap group. It would be equivalent to adding the `row_number()` by group after the `collect()`, but saves some typing.

```{r}
collect(x, id = ".id")
```

`original_id` tacks on the original row of the current bootstrap observation. It is generally more useful than `id`, as it provides a way to link the bootstrap rows back to the original data.

```{r}
collect(x, original_id = ".original_id")
```

# summarise()

The motivation for this package was `summarise()`. It efficiently computes the summary results, only materializing the bootstrap rows as they are needed at the C++ level.

```{r}
summarise(x, mean_length = mean(Sepal.Length))
```

You can group by other columns before creating the virtual groups, and `bootstrapify()` will respect those extra groups in the `summarise()` call. Pay attention to how easy it is to go from a non-bootstrapped version to a bootstrapped version. It's just one extra line!

```{r}
# Non-bootstrapped
iris %>%
  group_by(Species) %>%
  summarise(
    mean_length_across_species = mean(Sepal.Length)
  )

# Bootstrapped
iris %>%
  group_by(Species) %>%
  bootstrapify(5) %>%
  summarise(
    mean_length_across_species = mean(Sepal.Length)
  )
```

# do()

While `dplyr::do()` is basically deprecated and has been replaced by `group_map()`, it still has its uses sometimes. Like `summarise()`, `do()` materializes the groups only when they are required. Here we run the same linear model on each bootstrapped set of data.

```{r}
do(x, model = lm(Sepal.Length ~ Sepal.Width, data = .))
```

# group_nest()

`group_nest()` will materialize the groups so that they become columns in the outer tibble after the nest has been performed.

```{r}
group_nest(x)
```

You can set `keep = TRUE` to include the groups in the inner tibbles as well.

```{r}
group_nest(x, keep = TRUE)$data[[1]]
```

# group_split()

`group_split()` allows you to materialize all of the bootstrap tibbles into separate tibbles, all bundled together into a list.

```{r}
group_split(x) %>% head(n = 3)
```

You can specify `keep = FALSE` if you never want to see the bootstrap columns.

```{r}
group_split(x, keep = FALSE) %>% head(n = 3)
```

# group_map()

`group_map()` is similar to `do()`, but (as of dplyr 0.8.0.1) always returns a data frame and gives you access to the non-group and group data separately.

```{r}
# Just show the first 2 rows of each bootstrap
group_map(x, ~head(.x, n = 2))

# As you iterate though each group, you have access to that
# group's metadata through `.y` if you need it.
group_map_group_data <- group_map(x, ~tibble(.g = list(.y)))

group_map_group_data

group_map_group_data$.g[[1]]
```

Like `do()`, it can be a convenient way to run multiple models as long as you return a data frame from each one.

```{r}
iris %>%
  group_by(Species) %>%
  group_map(~ broom::tidy(lm(Petal.Length ~ Sepal.Length, data = .x)))
```

# ungroup()

`ungroup()` will return the original tibble back to you, without materializing the virtual groups.

```{r}
ungroup(x)
```

# as_tibble()

Like `ungroup()`, you can get the original tibble back by converting it to one explicitly with `as_tibble()`.

```{r}
as_tibble(x)
```

# Other dplyr functions

Most other dplyr functions work by first calling `collect()`, and then passing off to the underlying dplyr implementation. This means you can use `mutate()` like so:

```{r}
mutate(x, mean = mean(Sepal.Length))
```

This doesn't really get you anything in terms of speed, but can be convenient as an automatic way to convert back to a tibble and keep going with your workflow.
